---
title: "Corrplots"
date: "Today"
output: pdf_document
---

# Libraries 
```{r Libraries,echo=FALSE,message = FALSE,warning = FALSE}
knitr::opts_chunk$set(echo=FALSE,message = FALSE,warning = FALSE)

{
  pacman::p_load(
    here,
    janitor,
    tidyverse,ggpubr,broom,
    cowplot,ggx,vegan,
    patchwork,
    RColorBrewer) 
  
  theme_set(theme_cowplot())
  chron_colours <- c("#7D3C98","#9B30FF",  "darkgreen", "#74C365", "#FF8C00","#FFD700")
  
  }
```

# Data
```{r}

rds_files <- c("Bot", "Corrs_Bot", "Corrs_Fauna", "D18O", "D18O_WSr", 
               "Fauna", "Filtered_Bot", "Filtered_D18O_Bot", "Filtered_D18O_Fauna", 
               "Filtered_Fauna", "Joined_D18O_Bot", "Joined_D18O_Fauna")


for (file in rds_files) {
  full_file_path <- here("data", "pre-formatted dataframes", paste0(file, ".rds"))

  assign(file, readRDS(full_file_path))
}

```

# Selected Scatterplots for Discussion
## Fauna
```{r}

## Bos

Joined_D18O_Fauna %>% 
  filter(grepl("bos", species)) %>%
  filter(env_data=="se_Winter") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_smooth(method = "lm",se = FALSE,col="grey40",linetype="dashed")+
  geom_point(aes(col=chronology),show.legend = FALSE,size=3)+
  stat_cor(
    aes(label = paste(
      after_stat(r.label), after_stat(p.label), sep = "~`,`~"
    )),
    # label.x = 0.1,
    # label.y = 0.9,
    p.accuracy = 0.001,
    r.accuracy = 0.01,
    size=6
  ) +
  scale_color_manual(values = chron_colours) + 
  labs(title="Bos sp. and Winter Standard Error", x=" ", y="Normalised NISP", col="")

ggsave(filename = here("Scatterplots",paste("Bos-W_SE",".png")),width=20,height=20,units = "cm",bg = "white")

## Cervidae

Joined_D18O_Fauna %>% 
  filter(grepl("cervidae", species)) %>%
  filter(env_data=="range_Winter") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_smooth(method = "lm",se = FALSE,col="grey40",linetype="dashed")+
  geom_point(aes(col=chronology),show.legend = FALSE,size=3)+
  stat_cor(
    aes(label = paste(
      after_stat(r.label), after_stat(p.label), sep = "~`,`~"
    )),
    # label.x = 0.1,
    # label.y = 0.9,
    p.accuracy = 0.001,
    r.accuracy = 0.01,
    size=6
  ) +
  scale_color_manual(values = chron_colours) + 
  labs(title="Cervidae and Winter Range", x="Temperature in ºC", y="Normalised NISP", col="")

ggsave(filename = here("Scatterplots",paste("Cervidae-W_range",".png")),width=20,height=20,units = "cm",bg = "white")

## Dogs
 
Joined_D18O_Fauna %>% 
  filter(grepl("dog", species)) %>%
  filter(env_data=="Winter_Temp") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_smooth(method = "lm",se = FALSE,col="grey40",linetype="dashed")+
  geom_point(aes(col=chronology),show.legend = FALSE,size=3)+
  stat_cor(
    aes(label = paste(
      after_stat(r.label), after_stat(p.label), sep = "~`,`~"
    )),
    # label.x = 0.1,
    # label.y = 0.9,
    p.accuracy = 0.001,
    r.accuracy = 0.01,
    size=6
  ) +
  scale_color_manual(values = chron_colours) + 
  labs(title="Dogs and Winter Temperature", x="Temperature in ºC", y="Normalised NISP", col="")

ggsave(filename = here("Scatterplots",paste("Dog-W_temp",".png")),width=20,height=20,units = "cm",bg = "white")

#€ Equids

Joined_D18O_Fauna %>% 
  filter(grepl("equid", species)) %>%
  filter(env_data=="range_Winter") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_smooth(method = "lm",se = FALSE,col="grey40",linetype="dashed")+
  geom_point(aes(col=chronology),show.legend = FALSE,size=3)+
  stat_cor(
    aes(label = paste(
      after_stat(r.label), after_stat(p.label), sep = "~`,`~"
    )),
    # label.x = 0.1,
    # label.y = 0.9,
    p.accuracy = 0.001,
    r.accuracy = 0.01,
    size=6
  ) +
  scale_color_manual(values = chron_colours) + 
  labs(title="Equids and Winter Range", x="Temperature in ºC", y="Normalised NISP", col="")

ggsave(filename = here("Scatterplots",paste("Equids-W_range",".png")),width=20,height=20,units = "cm",bg = "white")

## Hare
Joined_D18O_Fauna %>% 
  filter(grepl("hare", species)) %>%
  filter(env_data=="Winter_Temp") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_smooth(method = "lm",se = FALSE,col="grey40",linetype="dashed")+
  geom_point(aes(col=chronology),show.legend = FALSE,size=3)+
  stat_cor(
    aes(label = paste(
      after_stat(r.label), after_stat(p.label), sep = "~`,`~"
    )),
    # label.x = 0.1,
    # label.y = 0.9,
    p.accuracy = 0.001,
    r.accuracy = 0.01,
    size=6
  ) +
  scale_color_manual(values = chron_colours) + 
  labs(title="Hare and Winter Temperature", x="Temperature in ºC", y="Normalised NISP", col="")

ggsave(filename = here("Scatterplots",paste("Hare-W_temp",".png")),width=20,height=20,units = "cm",bg = "white")


Joined_D18O_Fauna %>% 
  filter(grepl("hare", species)) %>%
  filter(env_data=="se_Winter") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_smooth(method = "lm",se = FALSE,col="grey40",linetype="dashed")+
  geom_point(aes(col=chronology),show.legend = FALSE,size=3)+
  stat_cor(
    aes(label = paste(
      after_stat(r.label), after_stat(p.label), sep = "~`,`~"
    )),
    # label.x = 0.1,
    # label.y = 0.9,
    p.accuracy = 0.001,
    r.accuracy = 0.01,
    size=6
  ) +
  scale_color_manual(values = chron_colours) + 
  labs(title="Hare and Winter Standard Error", x="", y="Normalised NISP", col="")

ggsave(filename = here("Scatterplots",paste("Hare-W_se",".png")),width=20,height=20,units = "cm",bg = "white")



Joined_D18O_Fauna %>% 
  filter(grepl("hare", species)) %>%
  filter(env_data=="se") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_smooth(method = "lm",se = FALSE,col="grey40",linetype="dashed")+
  geom_point(aes(col=chronology),show.legend = FALSE,size=3)+
  stat_cor(
    aes(label = paste(
      after_stat(r.label), after_stat(p.label), sep = "~`,`~"
    )),
    # label.x = 0.1,
    # label.y = 0.9,
    p.accuracy = 0.001,
    r.accuracy = 0.01,
    size=6
  ) +
  scale_color_manual(values = chron_colours) + 
  labs(title="Hare and Standard Error", x="", y="Normalised NISP", col="")

ggsave(filename = here("Scatterplots",paste("Hare-se",".png")),width=20,height=20,units = "cm",bg = "white")

## Large Ungulate

Joined_D18O_Fauna %>% 
  filter(grepl("large_ungulate", species)) %>%
  filter(env_data=="range_Winter") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_smooth(method = "lm",se = FALSE,col="grey40",linetype="dashed")+
  geom_point(aes(col=chronology),show.legend = FALSE,size=3)+
  stat_cor(
    aes(label = paste(
      after_stat(r.label), after_stat(p.label), sep = "~`,`~"
    )),
    # label.x = 0.1,
    # label.y = 0.9,
    p.accuracy = 0.001,
    r.accuracy = 0.01,
    size=6
  ) +
  scale_color_manual(values = chron_colours) + 
  labs(title="Large Ungulates and Winter Range", x="Temperature in ºC", y="Normalised NISP", col="")

ggsave(filename = here("Scatterplots",paste("L_Ung-W_range",".png")),width=20,height=20,units = "cm",bg = "white")

## Medium ungulate


Joined_D18O_Fauna %>% 
  filter(grepl("med_ungulate", species)) %>%
  filter(env_data=="Winter_Temp") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_smooth(method = "lm",se = FALSE,col="grey40",linetype="dashed")+
  geom_point(aes(col=chronology),show.legend = FALSE,size=3)+
  stat_cor(
    aes(label = paste(
      after_stat(r.label), after_stat(p.label), sep = "~`,`~"
    )),
    # label.x = 0.1,
    # label.y = 0.9,
    p.accuracy = 0.001,
    r.accuracy = 0.01,
    size=6
  ) +
  scale_color_manual(values = chron_colours) + 
  labs(title="Medium Ungulates and Winter Temperatures", x="Temperature in ºC", y="Normalised NISP", col="")

ggsave(filename = here("Scatterplots",paste("M_Ung-W_temp",".png")),width=20,height=20,units = "cm",bg = "white")

Joined_D18O_Fauna %>% 
  filter(grepl("med_ungulate", species)) %>%
  filter(env_data=="se") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_smooth(method = "lm",se = FALSE,col="grey40",linetype="dashed")+
  geom_point(aes(col=chronology),show.legend = FALSE,size=3)+
  stat_cor(
    aes(label = paste(
      after_stat(r.label), after_stat(p.label), sep = "~`,`~"
    )),
    # label.x = 0.1,
    # label.y = 0.9,
    p.accuracy = 0.001,
    r.accuracy = 0.01,
    size=6
  ) +
  scale_color_manual(values = chron_colours) + 
  labs(title="Medium Ungulates and Standard Error", x="", y="Normalised NISP", col="")

ggsave(filename = here("Scatterplots",paste("M_Ung-se",".png")),width=20,height=20,units = "cm",bg = "white")

Joined_D18O_Fauna %>% 
  filter(grepl("med_ungulate", species)) %>%
  filter(env_data=="range") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_smooth(method = "lm",se = FALSE,col="grey40",linetype="dashed")+
  geom_point(aes(col=chronology),show.legend = FALSE,size=3)+
  stat_cor(
    aes(label = paste(
      after_stat(r.label), after_stat(p.label), sep = "~`,`~"
    )),
    # label.x = 0.1,
    # label.y = 0.9,
    p.accuracy = 0.001,
    r.accuracy = 0.01,
    size=6
  ) +
  scale_color_manual(values = chron_colours) + 
  labs(title="Medium Ungulates and Temperature Range", x="Temperature in ºC", y="Normalised NISP", col="")

ggsave(filename = here("Scatterplots",paste("M_Ung-range",".png")),width=20,height=20,units = "cm",bg = "white")


## Small mammal
Joined_D18O_Fauna %>% 
  filter(grepl("small_mammal", species)) %>%
  filter(env_data=="Winter_Temp") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_smooth(method = "lm",se = FALSE,col="grey40",linetype="dashed")+
  geom_point(aes(col=chronology),show.legend = FALSE,size=3)+
  stat_cor(
    aes(label = paste(
      after_stat(r.label), after_stat(p.label), sep = "~`,`~"
    )),
    # label.x = 0.1,
    # label.y = 0.9,
    p.accuracy = 0.001,
    r.accuracy = 0.01,
    size=6
  ) +
  scale_color_manual(values = chron_colours) + 
  labs(title="Small Mammals and Winter Temperature", x="Temperature in ºC", y="Normalised NISP", col="")

ggsave(filename = here("Scatterplots",paste("Sm_Mam-Winter_temp",".png")),width=20,height=20,units = "cm",bg = "white")


## small ungulate

Joined_D18O_Fauna %>% 
  filter(grepl("small_ungulate", species)) %>%
  filter(env_data=="range_Winter") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_smooth(method = "lm",se = FALSE,col="grey40",linetype="dashed")+
  geom_point(aes(col=chronology),show.legend = FALSE,size=3)+
  stat_cor(
    aes(label = paste(
      after_stat(r.label), after_stat(p.label), sep = "~`,`~"
    )),
    # label.x = 0.1,
    # label.y = 0.9,
    p.accuracy = 0.001,
    r.accuracy = 0.01,
    size=6
  ) +
  scale_color_manual(values = chron_colours) + 
  labs(title="Small Ungulates and Winter Range", x="Temperature in ºC", y="Normalised NISP", col="")

ggsave(filename = here("Scatterplots",paste("Sm_U-Winter_range",".png")),width=20,height=20,units = "cm",bg = "white")
```




## Botanical

```{r}
## Graminae indeterminate
Joined_D18O_Bot %>% 
  filter(grepl("gramin", species)) %>%
  filter(env_data=="se_Summer") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_smooth(method = "lm",se = FALSE,col="grey40",linetype="dashed")+
  geom_point(aes(col=chronology),show.legend = FALSE,size=3)+
  stat_cor(
    aes(label = paste(
      after_stat(r.label), after_stat(p.label), sep = "~`,`~"
    )),
    # label.x = 0.1,
    # label.y = 0.9,
    p.accuracy = 0.001,
    r.accuracy = 0.01,
    size=6
  ) +
  scale_color_manual(values = chron_colours) + 
  labs(title="Graminae indet. and Summer Standard Error", x=" ", y="Normalised NISP", col="")

ggsave(filename = here("Scatterplots",paste("Gram_in-S_SE",".png")),width=20,height=20,units = "cm",bg = "white")

Joined_D18O_Bot %>% 
  filter(grepl("gramin", species)) %>%
  filter(env_data=="range_Summer") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_smooth(method = "lm",se = FALSE,col="grey40",linetype="dashed")+
  geom_point(aes(col=chronology),show.legend = FALSE,size=3)+
  stat_cor(
    aes(label = paste(
      after_stat(r.label), after_stat(p.label), sep = "~`,`~"
    )),
    # label.x = 0.1,
    # label.y = 0.9,
    p.accuracy = 0.001,
    r.accuracy = 0.01,
    size=6
  ) +
  scale_color_manual(values = chron_colours) + 
  labs(title="Graminae indet. and Summer Range", x="Temperature in ºC", y="Normalised NISP", col="")

ggsave(filename = here("Scatterplots",paste("Gram_in-S_range",".png")),width=20,height=20,units = "cm",bg = "white")

# Medicago

Joined_D18O_Bot %>% 
  filter(grepl("medic", species)) %>%
  filter(env_data=="se_Summer") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_smooth(method = "lm",se = FALSE,col="grey40",linetype="dashed")+
  geom_point(aes(col=chronology),show.legend = FALSE,size=3)+
  stat_cor(
    aes(label = paste(
      after_stat(r.label), after_stat(p.label), sep = "~`,`~"
    )),
    # label.x = 0.1,
    # label.y = 0.9,
    p.accuracy = 0.001,
    r.accuracy = 0.01,
    size=6
  ) +
  scale_color_manual(values = chron_colours) + 
  labs(title="Medicago sp. and Summer Standard Error", x="", y="Normalised NISP", col="")

ggsave(filename = here("Scatterplots",paste("Medic-S_se",".png")),width=20,height=20,units = "cm",bg = "white")



Joined_D18O_Bot %>% 
  filter(grepl("medic", species)) %>%
  filter(env_data=="range_Summer") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_smooth(method = "lm",se = FALSE,col="grey40",linetype="dashed")+
  geom_point(aes(col=chronology),show.legend = FALSE,size=3)+
  stat_cor(
    aes(label = paste(
      after_stat(r.label), after_stat(p.label), sep = "~`,`~"
    )),
    # label.x = 0.1,
    # label.y = 0.9,
    p.accuracy = 0.001,
    r.accuracy = 0.01,
    size=6
  ) +
  scale_color_manual(values = chron_colours) + 
  labs(title="Medicago sp. and Summer Range", x="Temperature in ºC", y="Normalised NISP", col="")

ggsave(filename = here("Scatterplots",paste("Medic-S_se",".png")),width=20,height=20,units = "cm",bg = "white")

## Pistacia

Joined_D18O_Bot %>% 
  filter(grepl("pist", species)) %>%
  filter(env_data=="se_Summer") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_smooth(method = "lm",se = FALSE,col="grey40",linetype="dashed")+
  geom_point(aes(col=chronology),show.legend = FALSE,size=3)+
  stat_cor(
    aes(label = paste(
      after_stat(r.label), after_stat(p.label), sep = "~`,`~"
    )),
    # label.x = 0.1,
    # label.y = 0.9,
    p.accuracy = 0.001,
    r.accuracy = 0.01,
    size=6
  ) +
  scale_color_manual(values = chron_colours) + 
  labs(title="Pistacia cf. lentiscus sp. and Summer Standard Error", x="", y="Normalised NISP", col="")

ggsave(filename = here("Scatterplots",paste("Pist-S_se",".png")),width=20,height=20,units = "cm",bg = "white")



Joined_D18O_Bot %>% 
  filter(grepl("pist", species)) %>%
  filter(env_data=="range_Summer") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_smooth(method = "lm",se = FALSE,col="grey40",linetype="dashed")+
  geom_point(aes(col=chronology),show.legend = FALSE,size=3)+
  stat_cor(
    aes(label = paste(
      after_stat(r.label), after_stat(p.label), sep = "~`,`~"
    )),
    # label.x = 0.1,
    # label.y = 0.9,
    p.accuracy = 0.001,
    r.accuracy = 0.01,
    size=6
  ) +
  scale_color_manual(values = chron_colours) + 
  labs(title="Pistacia cf. lentiscus sp. and Summer Range", x="Temperature in ºC", y="Normalised NISP", col="")

ggsave(filename = here("Scatterplots",paste("Pist-S_range",".png")),width=20,height=20,units = "cm",bg = "white")


## Triticum sp.



Joined_D18O_Bot %>% 
  filter(grepl("triti", species)) %>%
  filter(env_data=="range_Winter") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_smooth(method = "lm",se = FALSE,col="grey40",linetype="dashed")+
  geom_point(aes(col=chronology),show.legend = FALSE,size=3)+
  stat_cor(
    aes(label = paste(
      after_stat(r.label), after_stat(p.label), sep = "~`,`~"
    )),
    # label.x = 0.1,
    # label.y = 0.9,
    p.accuracy = 0.001,
    r.accuracy = 0.01,
    size=6
  ) +
  scale_color_manual(values = chron_colours) + 
  labs(title="Triticum sp. and Winter Range", x="Temperature in ºC", y="Normalised NISP", col="")

ggsave(filename = here("Scatterplots",paste("Triti-W_range",".png")),width=20,height=20,units = "cm",bg = "white")


Joined_D18O_Bot %>% 
  filter(grepl("triti", species)) %>%
  filter(env_data=="se_Winter") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_smooth(method = "lm",se = FALSE,col="grey40",linetype="dashed")+
  geom_point(aes(col=chronology),show.legend = FALSE,size=3)+
  stat_cor(
    aes(label = paste(
      after_stat(r.label), after_stat(p.label), sep = "~`,`~"
    )),
    # label.x = 0.1,
    # label.y = 0.9,
    p.accuracy = 0.001,
    r.accuracy = 0.01,
    size=6
  ) +
  scale_color_manual(values = chron_colours) + 
  labs(title="Triticum sp. and Winter Standard Error", x="", y="Normalised NISP", col="")

ggsave(filename = here("Scatterplots",paste("Triti-W_se",".png")),width=20,height=20,units = "cm",bg = "white")

Joined_D18O_Bot %>% 
  filter(grepl("triti", species)) %>%
  filter(env_data=="range") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_smooth(method = "lm",se = FALSE,col="grey40",linetype="dashed")+
  geom_point(aes(col=chronology),show.legend = FALSE,size=3)+
  stat_cor(
    aes(label = paste(
      after_stat(r.label), after_stat(p.label), sep = "~`,`~"
    )),
    # label.x = 0.1,
    # label.y = 0.9,
    p.accuracy = 0.001,
    r.accuracy = 0.01,
    size=6
  ) +
  scale_color_manual(values = chron_colours) + 
  labs(title="Triticum sp. and Temperature Range", x="Temperature in ºC", y="Normalised NISP", col="")

ggsave(filename = here("Scatterplots",paste("Triti-range",".png")),width=20,height=20,units = "cm",bg = "white")



## Triticum turgidum ssp. dicoccum

Joined_D18O_Bot %>% 
  filter(grepl("dicocc", species)) %>%
  filter(env_data=="range_Winter") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_smooth(method = "lm",se = FALSE,col="grey40",linetype="dashed")+
  geom_point(aes(col=chronology),show.legend = FALSE,size=3)+
  stat_cor(
    aes(label = paste(
      after_stat(r.label), after_stat(p.label), sep = "~`,`~"
    )),
    # label.x = 0.1,
    # label.y = 0.9,
    p.accuracy = 0.001,
    r.accuracy = 0.01,
    size=6
  ) +
  scale_color_manual(values = chron_colours) + 
  labs(title="Triticum turgidum ssp. dicoccum and Winter Range", x="Temperature in ºC", y="Normalised NISP", col="")

ggsave(filename = here("Scatterplots",paste("Triti_t_d-W_range",".png")),width=20,height=20,units = "cm",bg = "white")




```



# All Scatterplots
## Fauna

Based on the Corrs_Fauna data you can see which species are most correlated to which environmental proxies. Just change the name in the second line.
Based on the RDA data you can see which environmental factors might be most important.


### By Environmental data
```{r}
Joined_D18O_Fauna %>% 
  filter(env_data=="Winter_Temp") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_point(aes(col=chronology))+
  scale_color_manual(values = chron_colours) + 
  stat_cor(aes(label = paste(after_stat(r.label), after_stat(p.label), sep = "~`,`~")),label.x = 0.1, label.y = 0.9,p.accuracy = 0.001, r.accuracy = 0.01)+ 
    labs(title="Winter Minimum", x="Environmental Values", y="Normalised NISP", col="")+
  facet_wrap(~species,scales = "free")

Joined_D18O_Fauna %>% 
  filter(env_data=="range") %>% 
  ggplot()+
  aes(env_values,nisp)+
    geom_smooth(method = "lm",se = FALSE,col="grey40",linetype="dashed")+
  geom_point(aes(col=chronology))+
  scale_color_manual(values = chron_colours) + 
  stat_cor(aes(label = paste(after_stat(r.label), after_stat(p.label), sep = "~`,`~")),label.x = 0.1, label.y = 0.9,p.accuracy = 0.001, r.accuracy = 0.01)+ 
    labs(title="By annual range", x="Environmental Values", y="Normalised NISP", col="")+
  facet_wrap(~species,scales = "free")

Joined_D18O_Fauna %>% 
  filter(env_data=="se") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_point(aes(col=chronology))+
  scale_color_manual(values = chron_colours) + 
  stat_cor(aes(label = paste(after_stat(r.label), after_stat(p.label), sep = "~`,`~")),label.x = 0.1, label.y = 0.9,p.accuracy = 0.001, r.accuracy = 0.01)+ 
    labs(title="Variability within unit", x="Environmental Values", y="Normalised NISP", col="")+
  facet_wrap(~species,scales = "free")
```



### By Animal
```{r}
Joined_D18O_Fauna %>% 
  filter(species=="tuna") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_smooth(method = "lm",se = FALSE,col="grey40",linetype="dashed")+
  geom_point(aes(col=chronology))+
  stat_cor(aes(label = paste(after_stat(r.label), after_stat(p.label), sep = "~`,`~")),label.x = 0.1, label.y = 0.9,p.accuracy = 0.001, r.accuracy = 0.01)+ 
  scale_color_manual(values = chron_colours) + 
  labs(title="Tuna", x="Environmental Values", y="Normalised NISP", col="")+
  facet_wrap(~env_data,scales = "free")


```

## Botany 

Based on the Corrs_Bot data you can see which species are most correlated to which environmental proxies. Just change the name in the second line.
Based on the RDA data you can see which environmental factors might be most important.

### By Environmental data
```{r}

Joined_D18O_Bot %>% 
  filter(env_data=="Summer_Temp") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_point(aes(col=chronology))+
  scale_color_manual(values = chron_colours) + 
  stat_cor(aes(label = paste(after_stat(r.label), after_stat(p.label), sep = "~`,`~")),label.x = 0.1, label.y = 0.9,p.accuracy = 0.001, r.accuracy = 0.01)+ 
    labs(title="Summer maximum", x="Environmental Values", y="Normalised NISP", col="")+
  facet_wrap(~species,scales = "free")


Joined_D18O_Bot %>% 
  filter(env_data=="range_Winter") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_point(aes(col=chronology))+
  scale_color_manual(values = chron_colours) + 
  stat_cor(aes(label = paste(after_stat(r.label), after_stat(p.label), sep = "~`,`~")),label.x = 0.1, label.y = 0.9,p.accuracy = 0.001, r.accuracy = 0.01)+ 
    labs(title="By Winter range", x="Environmental Values", y="Normalised NISP", col="")+
  facet_wrap(~species,scales = "free")

Joined_D18O_Bot %>% 
  filter(env_data=="range") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_point(aes(col=chronology))+
  scale_color_manual(values = chron_colours) + 
  stat_cor(aes(label = paste(after_stat(r.label), after_stat(p.label), sep = "~`,`~")),label.x = 0.1, label.y = 0.9,p.accuracy = 0.001, r.accuracy = 0.01)+ 
    labs(title="By annual range", x="Environmental Values", y="Normalised NISP", col="")+
  facet_wrap(~species,scales = "free")

Joined_D18O_Bot %>% 
  filter(env_data=="se") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_point(aes(col=chronology))+
  scale_color_manual(values = chron_colours) +
  stat_cor(aes(label = paste(after_stat(r.label), after_stat(p.label), sep = "~`,`~")),label.x = 0.1, label.y = 0.9,p.accuracy = 0.001, r.accuracy = 0.01)+ 
    labs(title="Variability within unit", x="Environmental Values", y="Normalised NISP", col="")+
  facet_wrap(~species,scales = "free")

```



### By Plant
```{r}
Joined_D18O_Bot %>% 
  # filter(species=="triticum_turgidum_ssp_dicoccum") %>% 
  filter(grepl("tritic", species)) %>% 
  # filter(trenchunit!="FAS_83") %>%
  ggplot()+
  aes(env_values,nisp)+
  geom_smooth(method = "lm",se = FALSE,col="grey40",linetype="dashed")+
  geom_point(aes(col=chronology))+
  stat_cor(aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")),label.x = 0.1, label.y = 0.9,p.accuracy = 0.001, r.accuracy = 0.01)+ 
  scale_color_manual(values = chron_colours) +
  labs(title="Tritic", x="Environmental Values", y="Normalised NISP", col="")+
  facet_wrap(~env_data,scales = "free")

```
