---
title: "RDA Analysis Method"
author: "Niklas Hausmann"
date: "Today"
output: pdf_document
---

# Libraries 
```{r Libraries,echo=FALSE,message = FALSE,warning = FALSE}
knitr::opts_chunk$set(echo=FALSE,message = FALSE,warning = FALSE)

{
  pacman::p_load(
    here,
    janitor,
    tidyverse,ggpubr,broom,
    cowplot,ggx,vegan,
    patchwork,
    RColorBrewer) 
  
  theme_set(theme_cowplot())
  chron_colours <- c("#7D3C98","#9B30FF",  "darkgreen", "#74C365", "#FF8C00","#FFD700")
  
  }
```

# Data
## D18O
```{r}
D18O <- readxl::read_excel(here("data","arch_std_error.xls"), sheet = "Sheet1") %>% 
  as_tibble() %>% 
  clean_names() %>% 
  select(4,6:11,chronology) %>% 
  mutate(chronology=fct_relevel(chronology, c("FP","LM","UM","FM","FM/IN","IN","EN","EN-MN","MN","LN","FN"))) %>% 
  unite("trenchunit",trench,unit,sep="_") %>% 
  mutate(temp = 16.9-4.38*(d18o)+0.1*(d18o)*(d18o)) %>% 
  group_by(trenchunit,chronology) %>% 
  summarise(min=min(temp,na.rm = TRUE),
         max=max(temp,na.rm = TRUE),
         range=max-min,
         sd=sd(temp,na.rm = TRUE),
         se = sd(temp,na.rm = TRUE) / sqrt(n())
         ) %>% 
  ungroup() %>%
  arrange(trenchunit)


D18O_WSr <- readxl::read_excel(here("data","arch_std_error.xls"), sheet = "Sheet1") %>% 
  as_tibble() %>% 
  clean_names() %>% 
  select(4,6:11,chronology) %>% 
  unite("trenchunit",trench,unit,sep="_") %>%
  mutate(temp = 16.9-4.38*(d18o)+0.1*(d18o)*(d18o)) %>% 
  group_by(trenchunit,season) %>% 
  summarise(range=max(temp,na.rm = TRUE)-min(temp,na.rm = TRUE),
             se = sd(temp,na.rm = TRUE) / sqrt(n())
             ) %>% 
  ungroup() %>% 
  arrange(trenchunit) %>% 
  na.omit() %>% 
  pivot_wider(names_from = "season",values_from = c("range","se")) #%>% 
  #na.omit() # this deletes a lot of otherwise good data. it won't work with CCA and RDA with NAs but it will with the Scatterplots
  

```


## Fauna
```{r}
Fauna_FAN <- read.csv(here("data","Fauna_FAN_grouped_NISP.csv"),colClasses = c(unit="character")) %>% 
  clean_names() %>%
  mutate(trench="FAN")

Fauna_FAS <- read.csv(here("data","Fauna_FAS_grouped_NISP.csv"),colClasses = c(unit="character")) %>% 
  clean_names() %>% 
  mutate(trench="FAS")

Molluscs <- read.csv(here("data","Mollusc_data.csv"),colClasses = c(unit="character")) %>% 
  clean_names() %>% 
  unite("trenchunit",trench,unit,sep="_")




Fauna <- bind_rows(Fauna_FAN,Fauna_FAS) %>% 
  unite("trenchunit",trench,unit,sep="_") %>% 
  mutate(across(everything(), ~replace_na(.x, 0))) %>% 
  filter(if_any(where(is.numeric), ~ .x > 0)) %>% #remove empty rows
  arrange(trenchunit) %>% 
  left_join(Molluscs %>% rename(molluscs=nisp) %>% select(molluscs,trenchunit),by = "trenchunit")

rm(Fauna_FAN,Fauna_FAS,Molluscs)
```

## Botanical
```{r}
Bot <- read.csv(here("data","Botanical_data.csv")) %>% 
  select(-Column) %>% 
  filter(trench=="FAS"|trench=="FAN") %>% 
  unite("trenchunit",trench,unit,sep="_") %>% 
  arrange(trenchunit) %>% 
  mutate(zone = factor(zone,
                      levels = c("Ia",
                                 "Ib",
                                 "I/II",
                                 "I/III",
                                 "II",
                                 "II/III",
                                 "III",
                                 "IV",
                                 "IV/V",
                                 "V",
                                 "V/VI",
                                 "VI",
                                 "VI/VII",
                                 "VII",
                                 "VIIa",
                                 "VIIb"
                                 )
                    )
         ) %>% 
    replace_na(list(NISP=0)) %>% 
    pivot_wider(names_from = "species",values_from = "NISP",values_fn = {sum}) %>%  # at this step fragments and complete are summed
    select_if(~any(. != 0)) %>% #Removing rows and columns that are all 0s
  rowwise() %>% 
  filter(any(c_across(where(is.numeric)) != 0)) %>%
  ungroup() %>% 
  left_join(D18O %>% select(trenchunit,chronology),by="trenchunit") %>% 
  filter(chronology=="IN"|chronology=="EN"|chronology=="EN-MN"|chronology=="MN"|chronology=="LN"|chronology=="FN")
```

# Faunal analysis
### Filtered Fauna
```{r}
Filtered_Fauna <- Fauna %>%
  filter(trenchunit %in% D18O$trenchunit) %>%
  arrange(trenchunit) %>%
  filter(trenchunit!="FAN_123") %>%  # that one has only one isotope sample causing NAs
  # filter(trenchunit!="FAN_174") %>%  # that one has only onesummer isotope samples
  column_to_rownames("trenchunit") %>% 
  mutate(across(everything(), ~(.x - min(.x, na.rm = TRUE)) / 
                                  (max(.x, na.rm = TRUE) - min(.x, na.rm = TRUE))))

Filtered_D18O_Fauna <- D18O %>% 
  filter(trenchunit %in% Fauna$trenchunit) %>%
  mutate(
   Winter_Temp = min,#16.9-4.38*(min)+0.1*(min)*(min),
    Summer_Temp = max, #16.9-4.38*(max)+0.1*(max)*(max),
    range=Summer_Temp  - Winter_Temp,
    # Summer_Temp = (min - min(min, na.rm = TRUE)) / (max(min, na.rm = TRUE) - min(min, na.rm = TRUE)),
    # Winter_Temp = (max - min(max, na.rm = TRUE)) / (max(max, na.rm = TRUE) - min(max, na.rm = TRUE)),
  # sd = (sd - min(sd, na.rm = TRUE)) / (max(sd, na.rm = TRUE) - min(sd, na.rm = TRUE)),
    # se = (se - min(se, na.rm = TRUE)) / (max(se, na.rm = TRUE) - min(se, na.rm = TRUE)),
    # range_normalized = (range - min(range, na.rm = TRUE)) / (max(range, na.rm = TRUE) - min(range, na.rm = TRUE))
        ) %>% 
  select(-min,-max, -sd) %>% 
  arrange(trenchunit) %>%
  filter(trenchunit!="FAN_123") %>%  # that one has only one isotope sample causing NAs
  # filter(trenchunit!="FAN_174") %>%  # that one has only one isotope sample causing NAs
  column_to_rownames("trenchunit") %>% 
# Here I am filtering for Neolithic only, to exclude variability that might be due to larger climatic patterns at the early Holocene
      filter(chronology=="IN"|chronology=="EN"|chronology=="EN-MN"|chronology=="MN"|chronology=="LN"|chronology=="FN") %>% 
  select(-chronology)



```
###############################
DONT NEED TO CHANGE ANYTHING UP TO HERE 
##############################

RDA analysis will tell you whether and if yes how the variability (inertia) of the faunal or botanical data is correlated with or explained by the environmental data. You can then play around with the environmental factors using the equations or you can pick out the most important ones when running all factors and show in more detail how they relate to the faunal and botanical species using the scatterplots and correlation tests (R,R2 and p). 
Using ths CorrsFauna or CorrsBot dataframe you can look at what species are mostly affected by environmental change and also highlight those using scatterplots.

This should give you enough to discuss the impact of climatic variability or temperature extremes onto human subsistence at Franchthi.

## CCA - not as applicable as RDA as we have quantifiable environmental data
```{r}
# Now apply the cca function
All_CCA <- cca(Filtered_Fauna,Filtered_D18O_Fauna)
All_CCA
plot(All_CCA)

MinMax_CCA <- cca(Filtered_Fauna ~ Summer_Temp + Winter_Temp, data=Filtered_D18O_Fauna)
MinMax_CCA
plot(MinMax_CCA)
```

## RDA
```{r}
All_RDA <- rda(Filtered_Fauna,Filtered_D18O_Fauna)
All_RDA
plot(All_RDA)


MinMax_RDA <- rda(Filtered_Fauna ~ Summer_Temp + Winter_Temp, data=Filtered_D18O_Fauna)
MinMax_RDA
plot(MinMax_RDA)

SummerSE_RDA <- rda(Filtered_Fauna ~ se + Summer_Temp, data=Filtered_D18O_Fauna)
SummerSE_RDA
plot(SummerSE_RDA)

WinterSE_RDA <- rda(Filtered_Fauna ~ se + Winter_Temp, data=Filtered_D18O_Fauna)
WinterSE_RDA
plot(WinterSE_RDA)

```


# Botanical analysis
## Filtering Bot Data
```{r}
Filtered_Bot <- Bot %>% 
 filter(trenchunit %in% D18O$trenchunit) %>% 
  filter(trenchunit!="FAN_123") %>%  # that one has only one isotope sample causing NAs
  filter(trenchunit!="FAN_174") %>%  # that one has only one isotope sample causing NAs
  filter(trenchunit!="FAS_128") %>%  # that one has only one isotope sample causing NAs
  filter(trenchunit!="FAN_127") %>%  # that one has only one summer isotope samples
  column_to_rownames("trenchunit") %>% 
  select(-zone,-total_flot,-chronology) %>%
  mutate(across(-liters, ~./liters)) %>%  # this divides the remains by volume of the sample
  select(-liters) %>% 
  mutate(across(everything(), ~ if (all(is.na(.x))) NA else {  # this normalises the data to be more comparable
    min_val <- min(.x, na.rm = TRUE)
    max_val <- max(.x, na.rm = TRUE)
    if (min_val == max_val) 0 else (.x - min_val) / (max_val - min_val)
  })) %>% 
    select(where(~!all(unique(.) %in% c(0, 1))))  # This selects columns that don't have only 0 and 1 as their unique values


Filtered_D18O_Bot <- D18O %>% 
  filter(trenchunit %in% Bot$trenchunit) %>%
  mutate(
    Summer_Temp = 16.9-4.38*(min)+0.1*(min)*(min),
    Winter_Temp = 16.9-4.38*(max)+0.1*(max)*(max),
    range=Summer_Temp  - Winter_Temp,
    # Summer_Temp = (min - min(min, na.rm = TRUE)) / (max(min, na.rm = TRUE) - min(min, na.rm = TRUE)),
    # Winter_Temp = (max - min(max, na.rm = TRUE)) / (max(max, na.rm = TRUE) - min(max, na.rm = TRUE)),
  # sd = (sd - min(sd, na.rm = TRUE)) / (max(sd, na.rm = TRUE) - min(sd, na.rm = TRUE)),
    # se = (se - min(se, na.rm = TRUE)) / (max(se, na.rm = TRUE) - min(se, na.rm = TRUE)),
    # range_normalized = (range - min(range, na.rm = TRUE)) / (max(range, na.rm = TRUE) - min(range, na.rm = TRUE))
        ) %>% 
  select(-min,-max, -sd) %>% 
  arrange(trenchunit) %>%
  filter(trenchunit!="FAN_123") %>%  # that one has only one isotope sample causing NAs
  filter(trenchunit!="FAN_174") %>%  # that one has only one isotope sample causing NAs
  filter(trenchunit!="FAS_128") %>%  # that one has only one isotope sample causing NAs
  filter(trenchunit!="FAN_127") %>%  # that one has only one summer isotope samples
  column_to_rownames("trenchunit") %>% 
  select(-chronology)


```


## CCA - not as applicable as RDA as we have quantifiable environmental data
```{r}
# Now apply the cca function
All_CCA <- cca(Filtered_Bot,Filtered_D18O_Bot)
All_CCA
plot(All_CCA)

MinMax_CCA <- cca(Filtered_Bot ~ Summer_Temp + Winter_Temp, data=Filtered_D18O_Bot)
MinMax_CCA
plot(MinMax_CCA)
```

## RDA
```{r}
All_RDA <- rda(Filtered_Bot,Filtered_D18O_Bot)
All_RDA
plot(All_RDA)


MinMax_RDA <- rda(Filtered_Bot ~ Summer_Temp + Winter_Temp, data=Filtered_D18O_Bot)
MinMax_RDA
plot(MinMax_RDA)

SummerSE_RDA <- rda(Filtered_Bot ~ se + Summer_Temp, data=Filtered_D18O_Bot)
SummerSE_RDA
plot(SummerSE_RDA)

WinterSE_RDA <- rda(Filtered_Bot ~ se + Winter_Temp, data=Filtered_D18O_Bot)
WinterSE_RDA
plot(WinterSE_RDA)


```


# Scatterplots
## Fauna
### Fauna Data wrangle
```{r}
Joined_D18O_Fauna <- cbind(Filtered_D18O_Fauna,Filtered_Fauna) %>% 
  mutate(trenchunit=rownames(.)) %>% 
  left_join(D18O_WSr,by="trenchunit") %>% #adding the incomplete seasonal data
  select(trenchunit,range_Summer,range_Winter,se_Summer,se_Winter,everything()) %>%
  pivot_longer(names_to = "species", values_to = "nisp",cols = 10:26) %>%  
  pivot_longer(names_to = "env_data", values_to = "env_values",cols = 2:9) %>% 
  left_join(D18O %>% select(chronology,trenchunit),by="trenchunit")
  
```


### Fauna Correlations
```{r}
# Split data by species and environmental data to perform correlations
Corrs_Fauna <- Joined_D18O_Fauna %>% 
  group_by(species, env_data) %>%
  nest() %>%
  mutate(correlation = map(data, ~ tryCatch({
    cor.test(.x$nisp, .x$env_values, method = "pearson")
  }, error = function(e) NULL))) %>%
  # Correctly filter out NULL results outside of mutate
  filter(map_lgl(correlation, ~ !is.null(.x))) %>%
  mutate(tidy_cor = map(correlation, broom::tidy)) %>%
  select(-data, -correlation) %>%
  unnest(tidy_cor) %>%
  mutate(species_env = paste(species, env_data, sep = "_vs_"),
         r.squared = estimate^2) %>%
  select(species_env, estimate, p.value, r.squared) # This makes many warning but it's ok. Lots of NAs are being created, that's why.

```

### Fauna Plots
Based on the Corrs_Fauna data you can see which species are most correlated to which environmental proxies. Just change the name in the second line.
Based on the RDA data you can see which environmental factors might be most important.


#### By Environmental data
```{r}
Joined_D18O_Fauna %>% 
  filter(env_data=="range") %>% 
  ggplot()+
  aes(env_values,nisp)+
    geom_smooth(method = "lm",se = FALSE,col="grey40",linetype="dashed")+
  geom_point(aes(col=chronology))+
  scale_color_manual(values = chron_colours) + 
  stat_cor(aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")),label.x = 0.1, label.y = 0.9,p.accuracy = 0.001, r.accuracy = 0.01)+ 
    labs(title="Winter Temp", x="Environmental Values", y="Normalised NISP", col="")+
  facet_wrap(~species,scales = "free")

Joined_D18O_Fauna %>% 
  filter(env_data=="Summer_Temp") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_point(aes(col=chronology))+
  scale_color_manual(values = chron_colours) + 
  stat_cor(aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")),label.x = 0.1, label.y = 0.9,p.accuracy = 0.001, r.accuracy = 0.01)+ 
    labs(title="Summer Temp", x="Environmental Values", y="Normalised NISP", col="")+
  facet_wrap(~species,scales = "free")

Joined_D18O_Fauna %>% 
  filter(env_data=="range") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_point(aes(col=chronology))+
  scale_color_manual(values = chron_colours) + 
  stat_cor(aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")),label.x = 0.1, label.y = 0.9,p.accuracy = 0.001, r.accuracy = 0.01)+ 
    labs(title="Temp Range", x="Environmental Values", y="Normalised NISP", col="")+
  facet_wrap(~species,scales = "free")
```


#### By Animal
```{r}
Joined_D18O_Fauna %>% 
  filter(species=="sus_scrofa") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_smooth(method = "lm",se = FALSE,col="grey40",linetype="dashed")+
  geom_point(aes(col=chronology))+
  stat_cor(aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")),label.x = 0.1, label.y = 0.9,p.accuracy = 0.001, r.accuracy = 0.01)+ 
  scale_color_manual(values = chron_colours) + 
  labs(title="Medium ungulate", x="Environmental Values", y="Normalised NISP", col="")+
  facet_wrap(~env_data,scales = "free")


```

## Botany 
### Botany Data wrangle
```{r}
Joined_D18O_Bot <- cbind(Filtered_D18O_Bot,Filtered_Bot) %>% 
  mutate(trenchunit=rownames(.)) %>% 
  pivot_longer(names_to = "species", values_to = "nisp",cols = alkanna_sp:vicia_lathyrus_sp) %>%  
  pivot_longer(names_to = "env_data", values_to = "env_values",cols = 1:4) %>% 
  left_join(D18O %>% select(chronology,trenchunit),by="trenchunit")
```

### Botany Correlations
```{r}
# Split data by species and environmental data to perform correlations. LOTS OF WARNINGS BUT OK
Corrs_Bot <- Joined_D18O_Bot %>%
    filter(chronology=="IN"|chronology=="EN"|chronology=="EN-MN"|chronology=="MN"|chronology=="LN"|chronology=="FN") %>% 
  group_by(species, env_data) %>%
  nest() %>%
  mutate(correlation = map(data, ~ tryCatch({
    cor.test(.x$nisp, .x$env_values, method = "pearson")
  }, error = function(e) NULL))) %>% # This makes many warning but it's ok. Lots of NAs are being created, that's why.
  # Correctly filter out NULL results outside of mutate
  filter(map_lgl(correlation, ~ !is.null(.x))) %>%
  mutate(tidy_cor = map(correlation, broom::tidy)) %>%
  select(-data, -correlation) %>%
  unnest(tidy_cor) %>%
  mutate(species_env = paste(species, env_data, sep = "_vs_"),
         r.squared = estimate^2) %>%
  select(species_env, estimate, p.value, r.squared) 

```

### Botany Plots
Based on the Corrs_Bot data you can see which species are most correlated to which environmental proxies. Just change the name in the second line.
Based on the RDA data you can see which environmental factors might be most important.

#### By Environmental data
```{r}
Joined_D18O_Bot %>% 
  filter(env_data=="Winter_Temp") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_point(aes(col=chronology))+
  # scale_color_manual(values = chron_colours) + 
  stat_cor(aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")),label.x = 0.1, label.y = 0.9,p.accuracy = 0.001, r.accuracy = 0.01)+ 
    labs(title="Winter Temp", x="Environmental Values", y="Normalised NISP", col="")+
  facet_wrap(~species,scales = "free")

Joined_D18O_Bot %>% 
  filter(env_data=="Summer_Temp") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_point(aes(col=chronology))+
  scale_color_manual(values = chron_colours) + 
  stat_cor(aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")),label.x = 0.1, label.y = 0.9,p.accuracy = 0.001, r.accuracy = 0.01)+ 
    labs(title="Summer Temp", x="Environmental Values", y="Normalised NISP", col="")+
  facet_wrap(~species,scales = "free")

Joined_D18O_Bot %>% 
  filter(env_data=="range") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_point(aes(col=chronology))+
  scale_color_manual(values = chron_colours) + 
  stat_cor(aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")),label.x = 0.1, label.y = 0.9,p.accuracy = 0.001, r.accuracy = 0.01)+ 
    labs(title="Temp Range", x="Environmental Values", y="Normalised NISP", col="")+
  facet_wrap(~species,scales = "free")
```



#### By Plant
```{r}
Joined_D18O_Bot %>% 
  filter(species=="alkanna_sp") %>% 
  ggplot()+
  aes(env_values,nisp)+
  geom_smooth(method = "lm",se = FALSE,col="grey40",linetype="dashed")+
  geom_point(aes(col=chronology))+
  stat_cor(aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")),label.x = 0.1, label.y = 0.9,p.accuracy = 0.001, r.accuracy = 0.01)+ 
  scale_color_manual(values = chron_colours) +
  labs(title="Hordeum sp.", x="Environmental Values", y="Normalised NISP", col="")+
  facet_wrap(~env_data,scales = "free")


```
